Reproducible Research - Project Assignment 1
=============================================

# Author  - Samrat Paul
# Date  - Nov,2014
# Coursera  - Reproducible Research 

# About
This project is for the Assignment-1 of Reproducible Research course in Coursera's Data Science 
specialization track. It contains the answer of a series of questions by using data collected from a FitBit.

# Synopsis

Goal of this project :

* loading and preprocessing data

* inputing missing values

* interpreting data to answer research questions

# DataSet

Data is downloaded from the course web site:

Dataset: Activity monitoring data.

The variables included in this dataset are:

* steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

* date: The date on which the measurement was taken in YYYY-MM-DD format

* interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset


# Loading and preprocessing the data

Codes related to loading the data after downloading it.


```{r}

activity.data <- read.csv("repdata-data-activity/activity.csv", colClasses = c("numeric", "character", "numeric"))
head(activity.data)

```

* Question 1. What is mean total number of steps taken per day?

* Using the Aggregate Function

```{r}

steps_by_day <- aggregate(steps ~ date, data = activity.data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")

```
* Finding the mean and median ??

rmean   <- mean(steps_by_day$steps)

```{r}

mean(steps_by_day$steps)

```

* Finding the Median ??

rmedian <- median(steps_by_day$steps)

```{r}

median(steps_by_day$steps)

```

* Question 2. What is the average daily activity pattern?

  Making a time series plot of the 5-minute interval (x-axis) and the average number 
of steps taken, averaged across all days (y-axis)

```{r}

time.series <- tapply(activity.data$steps, activity.data$interval, mean, na.rm = TRUE)

```

The Plot is

```{r}

plot(row.names(time.series), time.series, type = "l", xlab = "5-min interval", 
     ylab = "Average across all Days", main = "Average Number of Steps Taken", 
     col = "red")

```
 * Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max.interval <- which.max(time.series)
names(max.interval)

```

* Imputing missing values

* Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)


```{r}

activity.data.na <- sum(is.na(activity.data))
activity.data.na

```

* Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need 
to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 
5-minute interval,etc

Fist Na replaced by mean in 5 min interval

```{r}
StepsAverage <- aggregate(steps ~ interval, data = activity.data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(activity.data)) {
        obs <- activity.data[i, ]
        if (is.na(obs$steps)) {
                steps <- subset(StepsAverage, interval == obs$interval)$steps
                } else {
                        steps <- obs$steps
                        }
        fillNA <- c(fillNA, steps)
        }

```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}

new.activity <- activity.data
new.activity$steps <- fillNA
names(new.activity)

```

* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number      of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


```{r}

StepsTotal2 <- aggregate(steps ~ date, data = new.activity, sum, na.rm = TRUE)

```

The Histogram is

```{r}

hist(StepsTotal2$steps, main = "Total steps by day", xlab = "day", col = "red")

```

And the mean and median is

```{r}
mean(StepsTotal2$steps)

```

```{r}
median(StepsTotal2$steps)

```

  * Are there differences in activity patterns between weekdays and weekends?

```{r}
day <- weekdays(as.Date(new.activity$date))
daylevel <- vector()
for (i in 1:nrow(new.activity)) {
if (day[i] == "Saturday") {
daylevel[i] <- "Weekend"
} else if (day[i] == "Sunday") {
daylevel[i] <- "Weekend"
} else {
daylevel[i] <- "Weekday"
}
}
new.activity$daylevel <- daylevel
new.activity$daylevel <- factor(new.activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = new.activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")


```


```{r}

library(lattice)
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```



